Efficient and Effective Metric Space Indexing for Complete Record Linkage

### 2017-11-20 paper.tex
Update on Overleaf.

Metric space indexing (MSI) techniques require indexing structures to be constructed over a set of records that are then used to support comparison with another set. MSI data structures often support a rich set of similarity search operations over a set of values over a metric space (distances must support the triangle inequality). The operations supported include \textit{range search} where all objects $i$ within a distance $d$ for a query object $q$ are identified, i.e.\ $dist(i,q) \le d$) \textit{nearest-neighbour} (the closest value to q), and \textit{nearestN} (the N closest n objects to q). An overview of the many Metric space search techniques that are available may be found in \cite{Zezula2010}. In this paper we choose one metric space data structure the M-Tree \cite{paolociaccia2m} and investigate its efficacy for record linkage.

An M-tree is a balanced tree structure comprised of fixed sized nodes each have a fixed number of children. In addition to the children, each node contains a reference to an object being indexed, a pointer and the distance to its parent, and a radius (described below). The radius of an internal node called a \textit{routing-node} represents the  furthest distance to any of its children. Thus all the children may be visualised as being contained within a ball of radius \textit{r} from the parent. M-trees are grown from the roots upward and require dynamic tree rebalancing when set of children become full (of size \textit{M}).

>>>>

Metric space indexing (MSI) techniques provide indexing structures to
support the comparison of records in one set with those in another,
typically also offering similarity search operations. To create an MSI
data structure, it is necessary to define a distance measure between
records, with certain properties including the \emph{triangle
inequality}. Similarity search operations include
\textit{range-search(q,d)}, where all records within a distance $d$ of a
query record $q$ are identified; \textit{nearest-neighbour(q)},
returning the record with smallest distance to $q$; and
\textit{nearest-n(q,n)}, returning the $n$ closest records to $q$.
\cite{Zezula2010} gives an overview of MSI techniques.
Here we choose one metric space indexing structure, the M-tree
\cite{paolociaccia2m}, and investigate its efficacy for record linkage.

The M-tree is a balanced tree structure comprised of fixed sized nodes each have a fixed number of children. In addition to the children, each node contains a reference to an object being indexed, a pointer and the distance to its parent, and a radius (described below). The radius of an internal node called a \textit{routing-node} represents the  furthest distance to any of its children. Thus all the children may be visualised as being contained within a ball of radius \textit{r} from the parent. M-trees are grown from the roots upward and require dynamic tree rebalancing when set of children become full (of size \textit{M}).
